---
title: "KidsAIR: multicollinearity exposure analysis"
author: "Ethan Walker"
date: "Started 16 April 2020, Updated 16 April 2020"
output: pdf_document
header-includes:
    - \usepackage[labelformat=empty]{caption}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE, include = FALSE,
                      message = FALSE, warning = FALSE)
```

```{r, eval=TRUE, include=TRUE}
library(readxl)
library(naniar)
library(lubridate)
library(broom)
library(broom.mixed)
library(zoo)
library(lme4)
library(lmerTest)
library(mctest)
library(tidyverse)
library(knitr)
library(kableExtra)
```


```{r, eval=TRUE, include=TRUE}
# Load individual datasets

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

exposure_analysis_data_long <- 
  read_rds(paste0(file_path, "Output/exposure_analysis_data_long_new.rds")) 

exposure_analysis_data_medium <- 
  read_rds(paste0(file_path, "Output/exposure_analysis_data_medium.rds")) 

exposure_analysis_data_short <- 
  read_rds(paste0(file_path, "Output/exposure_analysis_data_short.rds"))

sums_rolling_pm_data <- 
  read_rds(paste0(file_path, "Output/sums_rolling_pm_data.rds"))

health_exposure_data_sampling_day <- 
  read_rds(paste0(file_path, "Output/health_exposure_data_sampling_day.rds"))
```

```{r, eval=TRUE, include=TRUE}
####### Format full data to use for analysis
analysis_data <- exposure_analysis_data_short %>% 
  filter(winter_id == 1 & treatment == "Placebo" & area == "WMT") %>% 
  filter(home != "WMA300") %>% 
  group_by(area, home) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  ungroup() %>% 
  # dichotomize variables by median value
  mutate(moisture_closest_2level = cut(moisture_closest, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE),
         moisture_winter_2level = cut(mean_moisture_winter, breaks = c(0, 10.5, 100),
                                 labels = c("<10.5", "10.5+"),
                                 right = FALSE)) %>% 
  mutate(temp_max_2level = cut(temp_indoor_max, breaks = c(0, 30, 100),
                                 labels = c("<30", "30+"),
                                 right = FALSE),
         temp_min_2level = cut(temp_indoor_min, breaks = c(0, 16.5, 100),
                                 labels = c("<16.5", "16.5+"),
                                 right = FALSE),
         rh_max_2level = cut(rh_indoor_max, breaks = c(0, 40, 100),
                                 labels = c("<40", "40+"),
                                 right = FALSE),
         rh_min_2level = cut(rh_indoor_min, breaks = c(0, 20, 100),
                                 labels = c("<20", "20+"),
                                 right = FALSE)) %>% 
  mutate(sums_mean_sampling_period_5 = sums_mean_sampling_period/5,
         temp_perc_25_2level = cut(temp_perc_25, breaks = c(0, 50, 100),
                                 labels = c("<50", "50+"),
                                 right = FALSE),
         temp_perc_27_2level = cut(temp_perc_27, breaks = c(0, 35, 100),
                                 labels = c("<35", "35+"),
                                 right = FALSE),
         temp_perc_30_2level = cut(temp_perc_30, breaks = c(0, 10, 100),
                                 labels = c("<10", "10+"),
                                 right = FALSE)) %>% 
  mutate(home_act_door_2level = cut(home_act_door_sum, breaks = c(0, 1, 7),
                                 labels = c("<1", "1+"),
                                 right = FALSE),
         home_act_smoke_2level = cut(home_act_smoking_sum, breaks = c(0, 1, 7),
                                 labels = c("<1", "1+"),
                                 right = FALSE),
         home_act_windows_2level = cut(home_act_windows_sum, breaks = c(0, 1, 7),
                                 labels = c("<1", "1+"),
                                 right = FALSE),
         home_act_sweep_2level = cut(home_act_sweep_sum, breaks = c(0, 3, 7),
                                 labels = c("<3", "3+"),
                                 right = FALSE))
```


# Run tests for multicollinearity between different covariates
Goal is to check for vars that have a lot of overlap and eliminate some from
consideration to include in models

First, use `mctest` package to check for multicollinearity among similar variables

https://cran.r-project.org/web/packages/mctest/mctest.pdf 

Notes on multicollinearity:

https://datascienceplus.com/multicollinearity-in-r/ 

https://www.theanalysisfactor.com/eight-ways-to-detect-multicollinearity/ 

```{r, eval=TRUE, include=TRUE}
# Filter data for specific variables; select pm variable
pm_mean_sampling_period <- analysis_data %>% 
  select(pm_mean_sampling_period, 
         income_3level,
         residents_smoke,
         home_floors_2level, 
         home_sqft_2level,
         home_pets_2level,
         stove_age_3level, 
         chimney_clean_3level, 
         wood_collect_2level,
         moisture_closest_2level,
         sums_events_sampling_period,
         mean_temp_sampling_period_5,
         stove_grade_3level,
         home_fireplace) %>% 
  #mutate(area = factor(area)) %>% 
  mutate_all(as.numeric) %>% 
  filter(!is.na(pm_mean_sampling_period)) %>% 
  #na.exclude() %>% 
  select(pm_mean_sampling_period)

# Filter data for specific variables; select predictor variables
other_vars <- analysis_data %>% 
  select(pm_mean_sampling_period, 
         income_3level,
         residents_smoke,
         home_floors_2level, 
         home_sqft_2level,
         home_pets_2level,
         stove_age_3level, 
         chimney_clean_3level, 
         wood_collect_2level,
         moisture_closest_2level,
         sums_events_sampling_period,
         mean_temp_sampling_period_5,
         stove_grade_3level,
         home_fireplace) %>% 
  #mutate(area = factor(area)) %>% 
  mutate_all(as.numeric) %>% 
  filter(!is.na(pm_mean_sampling_period)) %>% 
  #na.exclude() %>% 
  select(-pm_mean_sampling_period)


# Use `omcdiag` to assess overall multicollinearity
#omcdiag(other_vars, pm_mean_sampling_period)


# Use `imcdiag` to assess multicollinearity between individual variables
#imcdiag(other_vars, pm_mean_sampling_period)

# Specify a method to assess
imcdiag(other_vars, log(pm_mean_sampling_period), method = "VIF")
```

Initially assessed the following variables for collinearity, based on simple
associations with mean PM over the sampling period:
         pm_mean_sampling_period, income_3level, area,
         residents_smoke,
         home_floors_2level, home_sqft_2level,
         home_pets_2level,
         stove_age_3level, chimney_clean_3level, 
         stove_cert, 
         wood_collect_2level,
         moisture_closest_2level,
         sums_events_sampling_period,
         mean_temp_sampling_period_5,
         amb_pm_sampling_period_5,
         home_act_smoke_2level, 
         stove_grade_3level,
         race_parent_2level, residents_under5_2level,
         home_type_2level, 
         home_fireplace 

The goal is to remove variables that have overlap, and parse the covariates
down to a reasonable number that a best-subsets selection method can handle.

This is the first sign that `area` has overlap with the other covariates that
will be difficult to deal with. Removing `area` lowers the VIF for several 
of the variables. 

Based on results and potential overlap, the following vars were removed:
         race_parent_2level, home_damp, home_mold, stove_age_3level,
         temp_max_5, home_act_gas, home_type_2level, 
         home_windows_2level, home_fireplace 
         
With the above variables removed, the VIF is at 5 or below for all variables.
There is still a question of how to handle the "area" variable, but I will
run best-subsets next and go from there.

Initial variables to use for model selection (14):
Demographics: area, income_3level, residents_smoke
Home Characteristics: home_floors_2level, home_pets_2level, stove_cert, 
                      chimney_clean_3level, wood_collect_2level, 
Home Activity: home_act_door_2level, home_act_smoke_2level
Others: sums_events_sampling_period, mean_temp_sampling_period_5, 
        stove_grade_3level, moisture_closest_2level, amb_pm_sampling_period_5,
         
Run sensitivity analyses to see if adding other variables back in makes a difference.

Can also try individual chi-square tests and correlations b/w two vars:

```{r}
test_data <- exposure_analysis_data_medium %>% 
  filter(winter_id == 1) %>% 
  filter(!is.na(pm_mean_sampling_day))

chisqu_results <- chisq.test(test_data$home_floors_2level, 
                             test_data$home_windows_2level)
chisqu_results



cor_results <- cor.test(test_data$temp_indoor_max, 
                        test_data$mean_temp)
cor_results
```

