---
title: "KidsAIR intervention questionnaires cleaning"
author: "Ethan Walker"
date: "Started 20 Jan 2021, Updated 20 Jan 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(readxl)
library(naniar)
library(lubridate)
library(zoo)

file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 
demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds")) 
```

# Load, format, and save Post Winter Questionnaire - filter homes
```{r}
##### WMT #####
# Load data
pwq_wmt <- read_csv(paste0(file_path, "Input/WMT/eow.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "WMT") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:helpful) %>% 
  ungroup() %>% 
  arrange(home_winter_id)


##### NN #####
# Load data
pwq_nn <- read_csv(paste0(file_path, "Input/NN/eow.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "NN") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:helpful) %>% 
  ungroup() %>% 
  arrange(home_winter_id)


##### AK #####
# Load data
pwq_ak <- read_csv(paste0(file_path, "Input/AK/eow.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "AK") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:helpful) %>% 
  ungroup() %>% 
  arrange(home_winter_id)


kids_pwq_filter <- rbind(pwq_wmt, pwq_nn, pwq_ak)

write_rds(kids_pwq_filter, paste0(file_path, "Output/kids_pwq_filter.rds"))
```

# Clean up all Post Winter Questionnaire - filter homes
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_pwq_filter <- read_rds(paste0(file_path, "Output/kids_pwq_filter.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 

kids_pwq_filter_clean <- kids_pwq_filter %>% 
  mutate(area = as.factor(area)) %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%
  filter(!is.na(treatment_assigned)) %>% 
  mutate(enteredby = as.factor(enteredby),
         fuse = as.factor(fuse),
         continuously = as.factor(continuously),
         reasonnoise = as.factor(reasonnoise),
         reasoncost = as.factor(reasoncost),
         reasonappearance = as.factor(reasonappearance),
         reasonspace = as.factor(reasonspace),
         reasoninconvenience = as.factor(reasoninconvenience),
         mostimp = as.factor(mostimp),
         helpful = as.factor(helpful)) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

summary(kids_pwq_filter_clean)

write_rds(kids_pwq_filter_clean, paste0(file_path, "Output/kids_pwq_filter_clean.rds"))
```

###############################################

# Load, format, and save Post Winter Questionnaire - education homes
```{r}
##### WMT #####
# Load data
pwq_wmt <- read_csv(paste0(file_path, "Input/WMT/eowe.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "WMT") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:comments) %>% 
  ungroup() %>% 
  arrange(home_winter_id)


##### NN #####
# Load data
pwq_nn <- read_csv(paste0(file_path, "Input/NN/eowe.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "NN") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:comments) %>% 
  ungroup() %>% 
  arrange(home_winter_id)


##### AK #####
# Load data
pwq_ak <- read_csv(paste0(file_path, "Input/AK/eowe.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "AK") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:comments) %>% 
  ungroup() %>% 
  arrange(home_winter_id)


kids_pwq_education <- rbind(pwq_wmt, pwq_nn, pwq_ak)

write_rds(kids_pwq_education, paste0(file_path, "Output/kids_pwq_education.rds"))
```

# Clean up all Post Winter Questionnaire - education homes
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_pwq_education <- read_rds(paste0(file_path, "Output/kids_pwq_education.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 

kids_pwq_education_clean <- kids_pwq_education %>% 
  mutate(area = as.factor(area)) %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%
  filter(!is.na(treatment_assigned)) %>% 
  mutate_if(is.character, as.factor) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

summary(kids_pwq_education_clean)

write_rds(kids_pwq_education_clean, paste0(file_path, "Output/kids_pwq_education_clean.rds"))
```

###############################################

# Load, format, and save KAB Questionnaire - filter homes
```{r}
##### WMT #####
# Load data
kab_wmt <- read_csv(paste0(file_path, "Input/WMT/kab.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "WMT") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:knowstove) %>% 
  ungroup() %>% 
  arrange(home_winter_id)


##### NN #####
# Load data
kab_nn <- read_csv(paste0(file_path, "Input/NN/kab.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "NN") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:knowstove) %>% 
  ungroup() %>% 
  arrange(home_winter_id)


##### AK #####
# Load data
kab_ak <- read_csv(paste0(file_path, "Input/AK/kab.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "AK") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:knowstove) %>% 
  ungroup() %>% 
  arrange(home_winter_id)


kids_kab_filter <- rbind(kab_wmt, kab_nn, kab_ak)

write_rds(kids_kab_filter, paste0(file_path, "Output/kids_kab_filter.rds"))
```

# Clean up all KAB Questionnaire - filter homes
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_kab_filter <- read_rds(paste0(file_path, "Output/kids_kab_filter.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 

kids_kab_filter_clean <- kids_kab_filter %>% 
  mutate(area = as.factor(area)) %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%
  filter(!is.na(treatment_assigned)) %>% 
  mutate_all(as.factor) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

summary(kids_kab_filter_clean)

write_rds(kids_kab_filter_clean, paste0(file_path, "Output/kids_kab_filter_clean.rds"))
```

###############################################

# Load, format, and save KAB Questionnaire - education homes
```{r}
##### WMT #####
# Load data
kab_wmt <- read_csv(paste0(file_path, "Input/WMT/kabe.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "WMT") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:knowstove) %>% 
  ungroup() %>% 
  arrange(home_winter_id)


##### NN #####
# Load data
kab_nn <- read_csv(paste0(file_path, "Input/NN/kabe.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "NN") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:knowstove) %>% 
  ungroup() %>% 
  arrange(home_winter_id)


##### AK #####
# Load data
kab_ak <- read_csv(paste0(file_path, "Input/AK/kabe.csv"), 
                                  na = c("NULL", "", "NA", -999)) %>% 
  mutate(home_winter_id = as.factor(HomeWinterID),
         area = "AK") %>% 
  rename_all(tolower) %>% 
  select(home_winter_id, area, enteredby:knowstove) %>% 
  ungroup() %>% 
  arrange(home_winter_id)


kids_kab_education <- rbind(kab_wmt, kab_nn, kab_ak)

write_rds(kids_kab_education, paste0(file_path, "Output/kids_kab_education.rds"))
```

# Clean up all KAB Questionnaire - education homes
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")
kids_kab_education <- read_rds(paste0(file_path, "Output/kids_kab_education.rds"))
kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 

kids_kab_education_clean <- kids_kab_education %>% 
  mutate(area = as.factor(area)) %>% 
  left_join(kids_linked_ids, by = c("area", "home_winter_id")) %>%
  filter(!is.na(treatment_assigned)) %>% 
  mutate_all(as.factor) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id)

summary(kids_kab_education_clean)

write_rds(kids_kab_education_clean, paste0(file_path, "Output/kids_kab_education_clean.rds"))
```

