---
title: "Kids SUMs Summary"
author: "Ethan Walker"
date: "Started 9 Jan 2020, Updated 7 Feb 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, 
                      include = TRUE, warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(readr)
library(readxl)
library(knitr)
library(lubridate)
jv_palette <- c("#330099","#CC0066","#FF6633", 
                 "#0099CC", "#FF9900","#CC6633",
                  "#FF3366", "#33CC99", "#33999")
```

```{r}
# Load data
sums_clean <- read_rds("Output/sums_clean.rds")
pm_clean <- read_rds("Output/pm_clean.rds")

```

# Summary stats by area (winter)
```{r}
sums_summary <- sums_clean %>% 
  # group by: area, treatment, dtid, day_of_week, Sampling_day
  #filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(temp_c_sums <= 0) %>% 
  group_by(area, home, ibid) %>% 
  summarize("n" = n(),
            "Mean Temp" = mean(temp_c_sums, na.rm = TRUE), 
            "SD Temp" = sd(temp_c_sums, na.rm = TRUE),
            "Min Temp" = min(temp_c_sums, na.rm = TRUE), 
            "Median Temp" = median(temp_c_sums, na.rm = TRUE),
            "Max Temp" = max(temp_c_sums, na.rm = TRUE)) %>% 
  select(-area) %>% 
  t() 

kable(sums_summary, col.names = c("Alaska", "Navajo", "Montana"), digits = 1, align = "c")


sums_summary <- sums_clean %>% 
  # group by: area, treatment, dtid, day_of_week, Sampling_day
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(temp_c_sums > 0) %>% 
  group_by(area) %>% 
  summarize("n" = n(),
            "Mean Temp" = mean(temp_c_sums, na.rm = TRUE), 
            "SD Temp" = sd(temp_c_sums, na.rm = TRUE),
            "Min Temp" = min(temp_c_sums, na.rm = TRUE), 
            "Median Temp" = median(temp_c_sums, na.rm = TRUE),
            "Max Temp" = max(temp_c_sums, na.rm = TRUE)) %>% 
  select(-area) %>% 
  t() 

kable(sums_summary, col.names = c("Alaska", "Navajo", "Montana"), digits = 1, align = "c")
```


# Closer look at negative temperatures
```{r}
total_homes <- sums_clean %>% 
  filter(winter_id == 1) %>% 
  count(home)

home_neg <- sums_clean %>% 
  filter(winter_id == 1) %>% 
  mutate(neg_temp = if_else(temp_c_sums < 0, 1, 0)) %>% 
  filter(neg_temp == 1) %>% 
  count(home) %>% 
  mutate(home_neg_temp = 1) %>% 
  select(home, home_neg_temp)

total_instruments <- sums_clean %>% 
  filter(winter_id == 1) %>% 
  count(ibid)

instrument_neg <- sums_clean %>% 
  filter(winter_id == 1) %>% 
  mutate(neg_temp = if_else(temp_c_sums < 0, 1, 0)) %>% 
  filter(neg_temp == 1) %>% 
  count(ibid) %>% 
  mutate(instrument_neg_temp = 1) %>% 
  select(ibid, instrument_neg_temp)

sums_neg_check <- sums_clean %>% 
  filter(winter_id == 1) %>% 
  left_join(home_neg, by = "home") %>% 
  left_join(instrument_neg, by = "ibid") %>% 
  mutate(home_neg_temp = if_else(is.na(home_neg_temp), 0, home_neg_temp),
         instrument_neg_temp = if_else(is.na(instrument_neg_temp), 0, instrument_neg_temp)) %>% 
  # group by both home_neg_temp and instrument_neg_temp to check summaries
  group_by(home_neg_temp) %>% 
  summarize("mean" = mean(temp_c_sums, na.rm = TRUE),
            "n" = n(),
            "min" = min(temp_c_sums, na.rm = TRUE),
            "median" = median(temp_c_sums, na.rm = TRUE),
            "max" = max(temp_c_sums, na.rm = TRUE))
sums_neg_check
```

The summaries above indicate that homes that had negative temperatures have similar
summary stats to those that did not have negative temps. The same is true for
the instruments used to measure temperature: instruments that had negative readings
have similar summary stats to those without negative readings. This gives some
reassurance that there are not any obvious trends associated with homes or
instruments that recorded negative temperatures during sampling.


# Boxplots for SUMs temp (C) by study area - Winter 1
```{r, fig.height=6, fig.width=5}
boxplots_kids_sums <- sums_clean %>%
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  ggplot(aes(area, temp_c_sums)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    ylim(-50, 100) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    labs(y = "Temperature (degrees C)") +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_sums
```

# Boxplots for SUMs temp (C) by study area - Winter 2
```{r, fig.height=6, fig.width=5}
boxplots_kids_sums <- sums_clean %>%
  filter(winter_id == 2) %>% 
  ggplot(aes(area, temp_c_sums)) +
    geom_boxplot(aes(size = 1), lwd = 1.2, colour = "black", 
               fatten = 1, outlier.size = 1.5, width = 0.35) +
    stat_summary(fun.y=mean, geom="point", shape=17, size=3, color="red") +
    theme_minimal() +
    ylim(-50, 100) +
    #geom_hline(yintercept = 50, color = "red", size = 1.2) +
    labs(y = "Temperature (degrees C)") +
    #labs(aes(x = "", y = paste("48-hour indoor PM "[2.5], " (", mu, g/m^3, ")"))) +
    theme(axis.text.x = element_text(size = 16, colour = "black"),
          axis.text.y = element_text(size = 16, colour = "black"),
          axis.title.y = element_text(size = 16,
                                      margin = margin(t = 0, r = 20, b = 0, l = 0)),
          axis.title.x = element_blank(),
          axis.line.x = element_line(colour = "black", size = 1.2), 
          axis.line.y = element_line(colour = "black", size = 1.2), 
          axis.ticks = element_blank())
boxplots_kids_sums
```

# Filter for homes that have PM data
```{r}
# 93 Winter 1 placebo homes have PM data
pm_homes_placebo <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  group_by(area, home) %>% 
  mutate(initial_obs = if_else(pm_datetime_new == first_datetime_new, 1, 0)) %>% 
  filter(initial_obs == 1) %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home, pm_datetime_new, pm) %>% 
  ungroup() %>% 
  arrange(area, home)

pm_homes_placebo_date <- pm_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  mutate(pm_home = "true") %>% 
  select(area, home, home_winter_id, pm_home, pm_datetime_new, pm) %>% 
  separate(pm_datetime_new, c("sample_date", "sample_time"), sep = " ", remove = FALSE) %>% 
  mutate(sample_date = ymd(sample_date)) %>% 
  group_by(area, home) %>% 
  distinct(sample_date, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home)

sums_filtered <- sums_clean %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo, by = c("area", "home", "home_winter_id")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home) 

sums_filtered_date <- sums_clean %>% 
  mutate(sample_date = date_sums,
         sample_date = ymd(sample_date)) %>% 
  # change to correct joining data of interest
  left_join(pm_homes_placebo_date, by = c("area", "home", "home_winter_id", "sample_date")) %>% 
  filter(pm_home == "true") %>% 
  filter(winter_id == 1) %>% 
  ungroup() %>% 
  arrange(area, home)

sums_winter1_placebo <- sums_clean %>% 
  filter(winter_id == 1) %>% 
  filter(treatment == "Placebo") %>% 
  ungroup() %>% 
  arrange(area, home)
```

# Function for calculating heating events using specified threshold temps
## Also specify the "gap" time between events
```{r}
thresholds_function <- function(data, temp, gap) {

thresholds <- data %>% 
  mutate(datetime_sums = ymd_hms(datetime_sums)) %>% 
  group_by(area, home, winter_id) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0),
         threshold_temp_event = if_else(threshold_temp == 1 & lag(threshold_temp) == 0, 
                                        "start", "Temp < threshold"),
         threshold_temp_event = if_else(is.na(threshold_temp_event), "start", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp == 1 & threshold_temp_event == "Temp < threshold",
                                      "Temp >= threshold", threshold_temp_event),
         threshold_temp_event = if_else(threshold_temp_event == "Temp >= threshold" &
                                      lead(threshold_temp_event) == "Temp < threshold", 
                                      "event_end", threshold_temp_event)) %>% 
  filter(threshold_temp_event == "event_end" | threshold_temp_event == "start") %>% 
  mutate(time_sums = as.character(time_sums),
         date_sums = as.character(date_sums)) %>%  
  spread(threshold_temp_event, datetime_sums) %>% 
  mutate(event_end = lead(event_end),
         event_length_min = (event_end - start),
         event_length_min = as.numeric(event_length_min/60)) %>% 
  filter(!is.na(event_length_min)) %>% 
  mutate(hrs_till_next_event = (lead(start)-event_end),
         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600,
         event_end2 = if_else(hrs_till_next_event <= gap, lead(event_end), event_end),
         event_end3 = if_else(hrs_till_next_event <= gap & lead(hrs_till_next_event) <= gap, 
                              lead(event_end, 2), event_end2),
         event_end4 = if_else(hrs_till_next_event <= gap & lead(hrs_till_next_event) <= gap
                              & lead(hrs_till_next_event, 2) <= gap, 
                              lead(event_end, 3), event_end3),
         event_end5 = if_else(hrs_till_next_event <= gap & lead(hrs_till_next_event) <= gap
                              & lead(hrs_till_next_event, 2) <= gap & lead(hrs_till_next_event, 3) <= gap, 
                              lead(event_end, 4), event_end4),
         event_end6 = if_else(hrs_till_next_event <= gap & lead(hrs_till_next_event) <= gap
                              & lead(hrs_till_next_event, 2) <= gap & lead(hrs_till_next_event, 3) <= gap
                              & lead(hrs_till_next_event, 4) <= gap, 
                              lead(event_end, 5), event_end5)) %>% 
  distinct(event_end6, .keep_all = TRUE) %>% 
  mutate(event_length_min = (event_end6 - start),
         event_length_min = as.numeric(event_length_min/60),
         hrs_till_next_event = (lead(start)-event_end),
         hrs_till_next_event = as.numeric(hrs_till_next_event)/3600) %>% 
  filter(!is.na(event_length_min)) %>% 
  group_by(area, home, winter_id, date_sums) %>% 
  summarize(n = n(), mean_time = mean(event_length_min))

thresholds_by_home <<- thresholds %>% 
  group_by(home) %>% 
  summarize("Mean events per day" = round(mean(n), digits = 2),
            "SD events per day" = round(sd(n), digits = 2),
            "Mean event length (min)" = round(mean(mean_time), digits = 0),
            "SD event length (min)" = round(sd(mean_time), digits = 2)) 
thresholds_by_home

thresholds_by_area <<- thresholds %>% 
  group_by(area) %>% 
  summarize("Mean events per day" = round(mean(n), digits = 2),
            "SD events per day" = round(sd(n), digits = 2),
            "Mean event length (min)" = round(mean(mean_time), digits = 0),
            "SD event length (min)" = round(sd(mean_time), digits = 2))
thresholds_by_area

}

thresholds_function(sums_filtered_date, 25, 1.5)
thresholds_function(sums_filtered_date, 27, 1.5)
thresholds_function(sums_filtered_date, 30, 1.5)
```


# Function to look at percentage of time spent above a threshold
```{r}
thresholds_percent <- function(data, temp) {

thresholds <- data %>% 
  group_by(area, home, winter_id, date_sums) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0)) %>% 
  summarize(percent_above = sum(threshold_temp)/n()*100) %>% 
  group_by(area) %>% 
  summarize("% time/day > threshold" = round(mean(percent_above), digits = 2))
thresholds

}

thresholds_percent(sums_filtered_date, 25)
thresholds_percent(sums_filtered_date, 27)
thresholds_percent(sums_filtered_date, 30)
```


# Define events by rate of change in temperature
```{r}
sums_temp_change <- sums_clean %>% 
  filter(winter_id == 1) %>% 
  group_by(area, home) %>% 
  mutate(time_diff = lead(datetime_sums) - datetime_sums) %>% 
  mutate(lead_temp = if_else(time_diff < 20, lead(temp_c_sums, 4), lead(temp_c_sums)),
         lead_temp = if_else(time_diff > 25, temp_c_sums, lead_temp),
         temp_diff = lead_temp - temp_c_sums,
         temp_diff_av = abs(lead_temp - temp_c_sums),
         temp_diff_check = if_else(temp_diff >= 5, 1, 0),
         heat_event = if_else(temp_diff_check == 1 & lag(temp_diff_check) == 1, 0,
                              temp_diff_check)) %>% 
  ungroup() %>% 
  arrange(area, home, datetime_sums)


summary_by_area <- sums_temp_change %>% 
  group_by(area, home, date_sums) %>% 
  summarize("events" = sum(heat_event, na.rm = TRUE)) %>% 
  group_by(area) %>% 
  summarize("Mean events" = mean(events),
            "SD events" = sd(events),
            "Min events" = min(events),
            "Median events" = median(events),
            "Max events" = max(events))
summary_by_area


temp_diff_by_area <- sums_temp_change %>% 
  mutate(time_diff = as.numeric(time_diff)) %>% 
  filter(time_diff >= 5 & time_diff < 30) %>% 
  group_by(area) %>% 
  summarize("mean temp diff" = mean(temp_diff_av, na.rm = TRUE),
            "mean time diff" = mean(time_diff, na.rm = TRUE))
temp_diff_by_area
```

